Her name was \sharkname, \sharknamefull. No one had ever used her first name, though. As a matter of fact, no one had ever dared to speak to her at all. If they did, they'd probably use \sharknameformal. But there would still be the other names, compulsively popping up in the mind of anyone unfortunate enough to meet her. The names were very much descriptive and seldom nice. `Lady in a grey dress' was one she could have liked, but then it went like `merciless queen,' `snapping shadow,' `fear from below,' `grey death,' \dots Or, simply, `the monster.'

Yet she was no monster. She was just a shark, fulfilling her important duty. Which was keeping her territory healthy. And that meant daily culling of animals that were too reckless, lazy or sick. The fact that this happened to result in a bit of carnage and a constant supply of fresh food was just a mildly pleasing bonus to \sharknameformal.

Despite her foul reputation, \sharknameformal was a particularly beautiful shark. Somehow slender, but far from the rather emaciated looks of her oceanic blue cousins. Her body, just over three meters long, spoke grace and power as she cruised along on her errands. Her back was grey and belly white, to hide her from wary eyes above and below \dots until she decides it's time to meet. Her tail was a little asymmetric, but in a very elegant way. Especially the long upper lobe looked very fancy, with its aggressive outlines. Her sickle-shaped dorsal fin was standing tall and proud, and the large pointed pectorals were tipped with a smooth gradient of darker shades. 

The first thing to note on her face were the eyes, attentive vertical slits of darkness in round pools of grey. Her broad snout sported two groove-like nose-flaps on each side. But the most imposing feature was certainly the mouth. It was curved in a perpetual frown, and full of sharp, serrated white teeth, means for delivering the communion of death to her chosen subjects.

\mybreak

The place \sharknameformal lived in was a little secluded corner of the vast ocean, a flooded bowl of a dormant volcano raising from the depths in the middle of nowhere. She was able to circumnavigate it in less than an hour at a very comfortable pace. The bottom was about sixty meters at its deepest. There were steep cliffs rising out of the water nearly everywhere, except of a small beach on the northern side and a breach in the southern wall, few hundred meters across. That was where the crater was invaded by the ocean not long after the volcano stopped spurting lava underwater and decided to become an island.

The bay was protected from ocean swells and currents. The water was calm and warm inside, and murky green with dense aggregations of phytoplankton. It looked much like a sleepy inland lake, blooming with unsavoury algae growth during a hot and heavy summer. 

There were surely prettier places in the ocean, but the bay was a much appreciated haven for many creatures small and large. The plankton was a staple for millions of tiny shrimps, bristly urchins, extravagant anemones and even occasional manta rays, gliding in on their huge wings from the open ocean. Hundreds of fish species preyed on the plankton-eaters and each other in turn. The bigger residents of the bay included curious octopuses, ponderous marine iguanas and playful fur seals.

Complex interactions between the plentiful life forms weaved the bay together; it ticked like a perfect organic clockwork. And its pace had been expertly controlled by \sharknameformal for a long time. She used to be the undisputed ruler of the depths and shallows alike. She was even respected and feared among the flocks of birds nesting high on the cliffs above the bay (they might have seemed out of her reach, and still \sharknameformal did sometimes manage to snatch a low-flying lava gull or storm petrel in a surprise breach).

Something, however, had started to change in the bay. For longer than she remembered, \sharknameformal could always be sure she'd have her daily share from the diverse menu the bay had on offer. But recently the food had become increasingly wary of her and thus harder to get hold of. She was still feared, but she was also being escaped much more often than before. She was even getting hungry sometimes, and wondered, with growing uneasiness, where this could all lead to.

\mybreak

Today \sharknameformal was feeling particularly blue. It was already afternoon and she hadn't eaten since the day before. And that was only a meager snack of few squids that got swept into the bay on the coming tide. 

\sharknameformal tried to hunt during the night. Plenty of tasty animals were out and about at that time. Octopuses, for instance. She had her fair share of the soft tentacly treats when she was a young lass, a bit too thin to take on larger prey. She was able to tackle much more ferocious (and nutritious) beasts now, but she was still fond of having a little squishy refreshment every now and then.

In the days back, octopuses were so easy to get. \sharknameformal just had to come and nudge them with her snout, as if inviting them for a game. The poor creatures were so curious that they usually started to crawl all over her head. With their obsession about exploring every nook and cranny, it was only a question of time before they started to stick their meaty tentacles to her mouth. Obviously, such a reckless inquisitiveness had to be punished and thus \sharknameformal devoured them without much ado then.

As time went, the octopuses had become trickier. It didn't start out so bad. Their growing mastery of mimicry might have made them slightly more difficult to spot, but \sharknameformal could find them even if they decided to resemble a piece of rock or an unsuspicious bump on the sandy bottom. No matter how perfect their visual disguise was, she could still sense their slightest move and even the faint triple-beats of their hearts if she happened to be close enough.

It got worse quickly, though. Recently the octopuses had become insults in the face of every honest and old-fashioned shark. They had grown far more careful and abandoned their suicidal curiosity. Their eyes seemed to be getting keener every day, so they could hardly be surprised. And even if one managed to somehow sneak upon them, they'd begun to use their mantles as means for incredibly fast propulsion. That often let them jet away to safety before \sharknameformal could snap her mouth shut on them.

Last night she thought she was getting lucky with an octopus, once after a long while. It was full moon. One of the chunkier representatives of the cephalopod genus, just about two nice mouthfuls, was gazing into the skies right in the middle of a large shallow patch of sandy bottom, far from any potential hiding holes. 

When she spotted this appetising scene on her nightly prowl, \sharknameformal did not hesitate and sprang into full speed. She was eager to punish the octopus for exposing itself so carelessly. 

Approaching fast, \sharknameformal was already opening her mouth. At the very last moment, however, something very puzzling happened. There was a quick dash of a blurry shape out of \sharknameformal' way and the octopus itself became strangely immaterial, as if made of dense smoke. Not thinking much about these new tricks, \sharknameformal sunk her teeth into the bulbous shape \dots and tore nothing out but empty coloured water. There was only a quickly dissolving cloud that used to look like an octopus a split second ago. 

The real octopus was darting away across the patch of sand, propelled by a quick successions of squirts from her jet engine. Enraged by this new treachery, \sharknameformal set on to pursuit the escape artist. The octopus was aiming for a rocky ridge extending from the shore into the bay. There must have been plenty of places to hide, but it was far, too far. The pace of the octopus jumps was fading already and the shark was still full of strength. 

\sharknameformal was accelerating hard. The octopus was very close to the rocks now. But it's jet engine seemed to have run out of juice. It could only swim with rhythmical contractions of the tentacles, hopelessly slow against the racing shark.

\sharknameformal was opening her mouth, this time sure her prey was doomed. She was painfully wrong again. Just a tiny moment before \sharknameformal would have bitten into the soft sinuous flesh, the octopus released a cloud of absolute darkness and fled on the last burst of its jet engine, smartly saved until the end of the race.

The outcome of this manoeuvre was quite pleasing for the octopus. It managed to find a narrow crack in the reef and squeeze itself in, hiding from any further pursuit. \sharknameformal was not happy at all, though. Still rushing at breakneck speed, she became utterly confused in the pitch-black slime the devious beast left behind. Not knowing where she's headed anymore, she started to slow down a tiny bit too late. When she emerged from the cloud, she could only watch a fast approaching rock wall and brace herself for a tough landing.

\mybreak

\sharknameformal nursed a nasty bump on her snout, a painful memory of the newest octopus invention. She decided to forget about this particular type of meal for a while. Surely there were more octopuses crawling about in the bay under the cover of the night, but there were also other, perhaps less deceitful customers she could pay a visit. Such as the fur seals.

Oh, sweet little fur seals, plump chunks of meat and blubber with flippery limbs and large cute eyes. They used to be such a treat when \sharknameformal first hunted in the bay on her own. Playing in the water for hours on end, they never cared too much about what was happening around. And if anything strange entered their territory, they foolishly tried to involve the intruder in their games. It had never crossed their minds that some visitors might be coming with not entirely friendly intentions (such as eating their hosts).

Only after many visits of \sharknameformal they stopped trying to play hide and seek with her. She always won; the hider was very well hidden soon, indeed. They also ceased to challenge her into a game of tag (back then, they seemed to think that her large size will make her slower than them \dots despite many bloody occasions that had proven such hopes false). 

Instead, they started to sprint to the safety of the shore as soon as they spotted her. But that was still fine by \sharknameformal. She could always lurk beneath their playing grounds, hidden in the murky green depths, and snatch the seals that ventured too far into the bay. When they saw her speeding from below, it was too late to retreat. It was not an endless playful feast anymore when the food stopped begging to be eaten and found out it can swim for its life, but she was still able to get a decent fatty supper whenever she wanted.

Not the last night, though. She knew the fur seals were hunting squids almost every night at the south-eastern wall of the bay. That would make them more alert than during their playtime, but the darkness also made the stealthy approach easier for \sharknameformal. She went deep, skimming the bottom on the way to make the hunters hunted.

Long before she could see their silhouettes darting in the moonlight above her, she sensed their moves \dots and their kills, as traces of squid blood dissolving in the ocean. She knew she was close. She swam in circles, ready to spring a trap as soon as a seal goes too far and deep on a squid chase. 

She could wait like that for the whole night, but this time she only hung there for few seconds before sensing the characteristic fluid movement of a seal nearby. Something was wrong, though. The seal was moving below and behind her. It must have been hiding motionless somewhere before she arrived. Soon she could even see the little animal, going fast right below her.

\sharknameformal didn't contemplate this unusual behaviour long and followed the seal. She was fast and it was a long way up. She was closing in on the poor guy quickly. But before she could grab the seal, there were more, springing up from holes and overhangs everywhere around her. The seals above stopped hunting and joined the ones on the bottom in a jittery dance around the shark. She didn't know which to follow, there were so many. Her instincts were telling her to target the closest and slowest prey, but she was failing at that miserably. The seals took turns crossing her path in wild patterns she couldn't ever hope to follow with the bulk of her body.

She was stalling inadvertently, shaking her head in confusion. And then the most shocking thing happened. She saw a seal approaching really close from behind, and before she could react, the little rascal went for her right eye with its tiny clawed front flipper. She jerked her whole body to avoid contact, but still received a painful bleeding jab just under the eye. At the same time, something raked against her left gills; another seal was coming for the other eye. 

It was hard to believe, but the cute little furry torpedoes were attacking \sharknameformal! Overwhelmed by this totally unexpected turn of events, she retreated to the depths, receiving few goodbye scratches in her face for a good measure. 
The injuries were not a big deal. Her rough skin could handle much harder tousling. What had brought her to the brink of depression was the shock of the prey fighting back. Such a thing was unheard of. If this was the way it was going to be in the bay now, she was in serious trouble.

\mybreak

Rattled by the humiliating encounter with the fur seals, \sharknameformal went as far as she could from their territory, to the sandy shallows on the other side of the bay. She did not want to see those cuddly little misfits for a long long time indeed. She stayed put until the sun went up. She needed to compose herself before trying luck elsewhere. 

In the morning, \sharknameformal regained her usual, simple and business-like attitude to life. She went exploring the eastern side of the bay, hoping for an easy breakfast to be had there. In that area, the bottom consisted of a dropping sequence of ledges and big boulders. And on these ledges, seeking shelter between the algae-covered stones and under numerous overhangs, marbled rays were often resting.

The rays were big, almost two meters across their round dark bodies mottled with lighter patches. They were longer still if one counted their thin tails. But they were also pretty sluggish during the day, and not too fast or agile even when fully alert. Hunting for them usually consisted of finding one that unwisely chose a space little too open for its rest. \sharknameformal could then just grab the all too accessible victim by a fin, drag it out into the open and quickly finish it off. 

The rays had recently started to be annoyingly successful in finding holes and overhangs way too narrow for \sharknameformal to reach in with her prying head. But once she spotted even a tiny patch of a juicy ray's body hiding in the cavelets and overhangs on the terraced bottom, she only had to wait nearby for them to move. They always did, sooner or later, and then they were lost.

As soon as she approached the ray quarters, she saw a massive one, the biggest she'd seen so far, almost fully exposed on a ledge. Her spirits were up, the embarrassing hunts of the night entirely forgotten for the moment. \sharknameformal didn't want to give the ray a chance to realise its mistake. She shifted into full attack speed. She aimed for delivering one mighty bite, crushing the ray's spine and ending the fight before it even begun. She was sure she was in for a well deserved breakfast. 

And she was wrong again. Just when she was about to open her mouth and sink her teeth into the succulent ray's body, she jerked back instead, blinded by a stabbing pain. The ray swam away from the reeling shark as if nothing happened, leisurely undulating the edges of its circular fins.

It was hard to say what exactly could have gone wrong. It all happened so damn fast. It seemed like the ray grew a thin barbed spike from its tail just when \sharknameformal was approaching. Then it raised this weirdly forked appendage, and right when \sharknameformal was about to bite, the ray used hurled the hard, pointy lance into her body.

The strike hit her deep in the back, just in front of her dorsal fin, but luckily it didn't do any major damage inside. Still, it was the most painful thing that'd ever happened to \sharknameformal, burning like thousand hells. It sent her squirming to deepest darkest spot of the bay, right on the bottom of the crater.

The water there was rarely disturbed by tides or currents. It was like a little lake within the ocean, less than a hundred meters across. The water was cooler there, and clearer than the turbid rest of the bowl up and around. Just what \sharknameformal needed now. Nursing the freshest piercing wound on the hump of her back, she descended into the pool and started swimming around ever so slowly, just to keep the soothing water flowing through her gills.

She'd have been crying if anyone had ever told her how. The pain was bad on its own, but so many failures in such a short time were really growing on her. An immaculate hunter and the undisputed ruler of the bay just a few days ago was now quickly turning into a lumbering spindle of useless muscles. It still had an impressive set of teeth on the pointy end, but these were no weapons to be feared anymore, much more an empty threat to be ridiculed.

\mybreak

For a long and dark while, \sharknameformal felt like she's done for. But the calm waters of the bottom lake worked wonders on her mood. The pain slightly subsided and what remained turned into new determination. She was not to be escaped and humiliated anymore. She was going for another hunt. 

Just to be on the safe side, though, she decided to go for the marine iguanas, some of the most sluggish swimmers that lived in the bay and were still worthy a bite (or two, in case of particularly large specimens). They usually lounged on land, basking on the black lava rocks near the shore during the day and snuggling to each other in cracks and crevices during the night. 

But they did go to the water to feed. Most of the sea lizards nibbled on the algae-covered reefs exposed by tide, but some more daring (and meatier) males also foraged underwater; they preferred to gorge themselves undisturbed by the others. 

These bold iguanas often went quite far and deep into the bay, and that was where \sharknameformal could lay an ambush. They could only go for so long, ripping the stubborn algae off the rocks with their bony jaws. The warmth of their bodies was quickly seeping into the water. Once they got too cold, they became slow and clumsy. They needed to go up and back to the shore to recharge on the sun. And that was \sharknameformal' time.

She didn't even have to swim particularly fast. She could just approach leisurely and grab the iguanas from the surface---no matter how hard they tried to snake through the water, they were still painfully slow for a shark.

Today it took \sharknameformal longer than usual to find a bite-worthy lizard. But she found one at last, scrapping the reefs scattered on the shallow bottom on the north-east side of the bay. It had its claws sunk into a mossy growth on a jagged rock that resembled a severed four-fingered hand half buried in sand. The shark kept a low profile and circled around, waiting for the iguana to spend all the warmth it brought from the surface.

The reptile took its time, greedily stripping half of the rock off its slimy algae veil. But then, at last, it went to the surface, to get a deep breath before swimming back to the shore. \sharknameformal didn't even wait for the iguana to reach the surface, interrupted her cautious circling and boldly entered the scene.

Before she could execute her trademark lazy snatch of the thorny piece of lizard meat, weird things started to happen again. It was like a recurrent nightmare, one seemingly harmless prankster replacing another to torture \sharknameformal.

When the iguana spotted the shark approaching, it didn't start to wiggle its body on the surface in a futile attempt to swim away. It took a quick breath instead and went right down to the bottom again before \sharknameformal could get it. Down there, it dug its hind legs into the craggy reef face, so conveniently cleared off all algae that could let go. The iguana's thorny face was following every move of \sharknameformal and its front legs were poised in a fighting stance, with their wide open claws glistening in the greenish light of underwater noon.

Puzzled by yet another unexpected twist, \sharknameformal tried few half-hearted moves on the stubborn warrior. But her spirits were down again and the lizard was aggressively parrying each of her wary approaches. Once she got a nasty scratch on her nose, she was done there. Before collecting any more wounds, she retreated, drifting randomly through the bay and sinking into despair.

\mybreak

The iguana farce was the last drop in \sharknameformal' cup of sorrow. Now she was hungry, sore all over her body and utterly dispirited. She could not see how this came to be. Her instincts, her speed and power, the brute force hunting strategies that seemed to work so beautifully just few days ago, all was useless now. 

Lost in the depths of her depression, listlessly roaming through the bay's water column, \sharknameformal unwittingly swam up towards the surface. And there her brooding was somewhat interrupted by yet another disturbing phenomenon she was not familiar with. Only this time the perpetrator was not an animal with a nasty surprise in its pocket, but the bay and nature itself. The normally bright midday sky went unusually dark, but there was still plenty of light diffused underwater. The bay's surface looked like a mirror from below, and there \sharknameformal beheld herself. 

She scanned through the fresh cuts and bruises, the only outcomes of her recent hunting adventures. The sight was pitiful. But she also felt something was rising deep within her, a wave of comprehension far beyond the simple life she'd led so far. 

Looking at all the scars on her still glorious body was bringing memories, something she'd never experienced before. All her fights won and lost were being replayed in her mind. And then she realised she can work with those memories. She knew what could have been done better, she could set on different paths in millions of alternative histories. In the myriads of possibilities, \sharknameformal saw many more victories and tasty meals than she could have imagined before. 

And thus it happened that the mindless killing machine started to think. She turned her back on the mirror and smiled for the first time in her life. And the whole bay shuddered, anxiously awaiting the hunts to come.

